Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd

```{r}
#| label: fig-explore-data
#| fig-cap: "Exploratory Analysis of Predictors and Candidate Support (pct)"
#| echo: false
#| warning: false
a <- ggplot(data=cleaned_data, aes(x=sample_size, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'sample size', y='pct', 
       title = 'pct vs sample size')

b <- ggplot(data=cleaned_data, aes(x=pollscore, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'pollscore', y='pct', 
       title = 'pct vs pollscore')

c <- ggplot(data=cleaned_data, aes(x=methodology, y=pct)) + 
  geom_boxplot() + 
  labs(x = 'methodology', y='pct', 
       title = 'pct vs methodology')

d <- ggplot(data=cleaned_data, aes(x=days_taken_from_election, y=pct)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'max gust speed', y='pct', 
       title = 'pct vs days taken from election')

e <- ggplot(data=cleaned_data, aes(x=state, y=pct)) + 
  geom_boxplot() + 
  labs(x = 'state', y='pct', 
       title = 'pct vs state')

grid.arrange(a, b, c, d, e, nrow=2)
```

@fig-explore-data explores the relationship between candidate support percentage (pct) and various predictors in the dataset: (a) sample size of polls, (b) pollscore (an indicator of poll reliability), (c) polling methodology, (d) days taken from the election, and (e) state.

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-explore-data in @sec-dataexplore, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct. Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

@@

>>>>

@@

We chose sample size, poll score, and days taken from the election as predictors for building the model based on the visual trends observed in the plots. As shown in @fig-pollsamplesize-canidate-pct, @fig-pollscore-candidate-pct and @fig-polldate-canidate-pct in @sec-data, The relationship between pct and sample size shows that larger sample sizes correspond to more stable percentages, indicating that sample size could significantly impact prediction accuracy. The plot of pct versus poll score reveals a negative trend, suggesting that poll scores are consistently influential in explaining variations in pct. Additionally, the pct versus days taken plot indicates that the proximity to the election affects the percentage, with a slight negative trend showing that time remaining has predictive value. These observed patterns imply that these variables can effectively capture key factors influencing pct and are suitable for model inclusion.

@@

## Sources of Bias In the Polls

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached. Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes. This trend may be worsened by decreased participation in polls as the election date approaches. As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates. Additionally, the withdrawal of a prominent third-party candidate, Robert F. Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic. However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the 538 poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could be strategically beneficial for the Harris campaign, providing a warning to avoid repeating previous electoral mistakes.
